---
title: "Coursera: Johns Hopkins Data Science Capstone - Milestone Report"
author: "Travis May"
output:
  html_document:
    highlight: textmate
    keep_md: yes
    theme: cosmo
    toc: yes
---
```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
library(RColorBrewer)
library(RWeka)
library(NLP)
library(tm)
library(SnowballC)
library(slam)
library(wordcloud)
library(stringr)
library(stringi)
library(googleVis)
```

```{r results='hide',eval=TRUE,echo=FALSE}
required.scripts  = c('get.R','load.R','clean.R','sample.R')
sapply(required.scripts, source, .GlobalEnv)
```
# Introduction:
The purpose of this milestone report is to indicate that I am sufficiently capable of handling large data sets within the R framework. The data set that I will use is a corpora of curated text from various blogs, news sites, and twitter posts, or 'tweets'. I will apply this data set to a text prediction algorithim which will be compiled into a Shiny application in the second half of the course.

The remaining sections of this report will first explain how I extract and clean the data, and then present the analysis of data required for successful implementation of the text prediction app.

# The Data:
The HC Corpora is sourced as provided by the course instructors. The data is located at http://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip. More detailed information about the corpora at http://www.corpora.heliohost.org/aboutcorpus.html.


## Data Processing
The corpora is available in multiple languages, but for the purposes of this project I will only use the files in US English.

### Download 
```{r, eval=FALSE, echo=TRUE}
fileURL <- "http://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"
download.file(fileURL, destfile = "Dataset.zip", method = "curl")
unlink(fileURL)
unzip("Dataset.zip")
```

### Read into Environment
I use a custom function to read in the data iteratively.The function also converts the data from 'latin1' encoding to 'UTF-8'.

#### The function:
```{r, eval=FALSE, echo=TRUE}
loader <- function(item) {

    if(tools::file_ext(item)=='txt') {

      # Import 'News' data in binary mode
      if(grepl('news',item)) {
        con <- file(paste('./final/en_US/',item,sep = '/'),open='rb')
        l <- readLines(con, encoding='UTF-8')
        close(con)
        rm(con)
      }
      # Import 'Twitter' and 'Blog' data in text mode
      else {
        l <- readLines(paste('./final/en_US/',item,sep = '/'),skipNul = TRUE,warn = FALSE, encoding = 'UTF-8')
      }
    }

    return(l)
}
```

#### Using the function on the downloaded files: 
```{r, eval=TRUE, echo=TRUE}
files = c(list.files('./final/en_US'))

for (i in files) {
  name = strsplit(i,"\\.")[[1]]
  assign(name[[2]],loader(i))
}
```

### Clean the data
Next, I use regular expressions to clean out any unusual characters that may affect my analysis.
```{r,eval=TRUE, echo=TRUE}
blogs <- iconv(blogs,from="latin1", to = "UTF-8", sub="")
blogs <- stri_replace_all_regex(blogs, "\u2019|`","'")
blogs <- stri_replace_all_regex(blogs, "\u201c|\u201d|u201f|``",'"')
news <- iconv(news,from="latin1", to = "UTF-8", sub="")
news <- stri_replace_all_regex(news, "\u2019|`","'")
news <- stri_replace_all_regex(news, "\u201c|\u201d|u201f|``",'"')
twitter <- iconv(twitter, from = "latin1", to = "UTF-8", sub="")
twitter <- stri_replace_all_regex(twitter, "\u2019|`","'")
twitter <- stri_replace_all_regex(twitter, "\u201c|\u201d|u201f|``",'"')
```

### Sample the data
Since the data is so large, I sample 10000 lines from each set of text and I save the file for later.
```{r,eval=TRUE,echo=TRUE}
sample_blogs   <- sample(blogs, 10000)
sample_news    <- sample(news, 10000)
sample_twitter <- sample(twitter, 10000)

sample <- c(sample_twitter,sample_news,sample_blogs)

writeLines(sample, "./sample.txt")
```

### Data Set Summary tatistical Table
Below is a table that gives a quick look at the data files compared with the sample data.
```{r,eval=TRUE,echo=FALSE}
blog_stats <- './final/en_US/en_US.blogs.txt'
news_stats <- './final/en_US/en_US.news.txt'
twitter_stats <- './final/en_US/en_US.twitter.txt'
sample_stats <- "./sample.txt"

fileSummary <- data.frame(
        fileName = c("Blogs","News","Twitter", "Aggregated Sample"),
        fileSize = c(round(blogs_stats, digits = 2), 
                     round(news_stats,digits = 2), 
                     round(twitter_stats, digits = 2),
                     round(sample_stats, digits = 2)),
        lineCount = c(blogsLength, newsLength, twitterLength, sampleLength),
        wordCount = c(blogsWords, newsWords, twitterWords, sampleLength)                
)

colnames(fileSummary) <- c("File Name", "File Size in Megabyte", "Line Count", "Word Count")

saveRDS(fileSummary, file = "./summary.Rda")

fileSummaryDF <- readRDS("./fileSummary.Rda")

```

```{r,eval=TRUE,echo=TRUE}
knitr::kable(head(fileSummaryDF, 10))
```

### Tokenization
I utilize the 'tm' package to transform the text, remove punctuation and stop words, and stem the document before loading the text into a matrix. This matrix is used to construct a word cloud.
```{r,eval=TRUE,echo=TRUE}
vs.s = VectorSource(sample)
s.corpus = VCorpus(vs.s)
s.corpus = tm_map(s.corpus, content_transformer(tolower))
s.corpus = tm_map(s.corpus, removePunctuation)
s.corpus = tm_map(s.corpus, removeNumbers)
s.corpus = tm_map(s.corpus, function(x) removeWords(x, stopwords("english")))
s.corpus = tm_map(s.corpus, stemDocument)

c_tdm    = TermDocumentMatrix(s.corpus)
c_tdm    = rollup(c_tdm, 2, na.rm=TRUE, FUN = sum)
c_tdm.m  = as.matrix(c_tdm)
c_tdm.v  = sort(rowSums(c_tdm.m),decreasing=TRUE)
c_tdm.d  = data.frame(word = names(c_tdm.v),freq=c_tdm.v)
```

```{r,eval=TRUE,echo=TRUE}
c_tdm    = TermDocumentMatrix(s.corpus)
c_tdm    = rollup(c_tdm, 2, na.rm=TRUE, FUN = sum)
c_tdm.m  = as.matrix(c_tdm)
c_tdm.v  = sort(rowSums(c_tdm.m),decreasing=TRUE)
c_tdm.d  = data.frame(word = names(c_tdm.v),freq=c_tdm.v)
table(c_tdm.d$freq)
pal2 = brewer.pal(8,"Dark2")
wordcloud(c_tdm.d$word,c_tdm.d$freq, scale=c(5,.3),min.freq=1000,
          max.words=Inf, random.order=FALSE, rot.per=.15, colors=pal2)
```

```{r, eval=FALSE, echo=FALSE}
## Saving the final corpus
saveRDS(s.corpus, file = "./finalCorpus.RDS")

final.corpus <- readRDS("./finalCorpus.RDS")

final.corpusDF <-data.frame(text=unlist(sapply(finalCorpus,`[`, "content")), 
                           stringsAsFactors = FALSE)
```

