---
title: "Coursera: Johns Hopkins Data Science Capstone - Milestone Report"
author: "Travis May"
output:
  html_document:
    highlight: textmate
    keep_md: yes
    theme: cosmo
    toc: yes
---

# Introduction:
The purpose of this milestone report is to indicate that I am sufficiently capable of handling large data sets within the R framework. The data set that I will use is a corpora of curated text from various blogs, news sites, and twitter posts, or 'tweets'. I will apply this data set to a text prediction algorithim which will be compiled into a Shiny application in the second half of the course.

The remaining sections of this report will explain how I extract, clean, and analyze the data required for the text prediction app which I will present at the end of the course.

# The Data:
The HC Corpora is sourced as provided by the course instructors. The data is located at http://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip. More detailed information about the corpora at http://www.corpora.heliohost.org/aboutcorpus.html.


## Data Processing
The corpora is available in multiple languages, but for the purposes of this project I will only use the files in US English.

### Download 
```{r, eval=FALSE, echo=TRUE}
fileURL <- "http://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"
download.file(fileURL, destfile = "Dataset.zip", method = "curl")
unlink(fileURL)
unzip("Dataset.zip")
```

### Read into Environment 
```{r, eval=FALSE, echo=FALSE}
blogs <- readLines("./final/en_US/en_US.blogs.txt", encoding = "UTF-8", skipNul=TRUE)
news <- readLines("./final/en_US/en_US.news.txt", encoding = "UTF-8", skipNul=TRUE)
twitter <- readLines("./final/en_US/en_US.twitter.txt", encoding = "UTF-8", skipNul=TRUE)
```